---
title: "Who am I?"
description: |
  The proverbial question of the ages.
---

```{r list, include = FALSE}
projects <- c("The influential impact of brand activism on subsequent consumer activism", "A conceptualization and examination of reactive disposal in the marketplace", "How a firm's adoption of LGBTQ+-inclusive policies affects stock performance", "Factors that drive the believability of 'clickbait' persuasive appeals")

d <- data.frame(projects)

library(knitr)
library(kableExtra)
options(knitr.table.format = "html")
```

<p style = "text-align:center;"><img src = "real_me.jpg" alt = "Logo" style = "width: 600px;"></p>

My name is **Andrew Edelblum**! 

I am a marketing Ph.D. student at the [University of Oregon](https://business.uoregon.edu/faculty/andrew-edelblum) and the social media manager for the [*Journal of Consumer Research*](http://www.ejcr.org/). I'm interested in consumer identity, activism, new media platforms, and all things Nintendo.

My current research projects involve:

```{r table}
kable(d, col.names = "Project") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))
```

Click [here](index.html) to check out some of my projects, datasets, and visualizations in greater detail!