---
title: "My Favorite Songs of All Time"
description: |
    I'm a big music fan, so I ranked my favorite tracks
author:
  - name: Andrew Edelblum
    url: https://twitter.com/andrewedelblum
date: 03-13-2019
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 7)

library(dplyr)
library(tidyverse)
library(ggplot2)
library(janitor)
library(rio)
library(knitr)
library(plotly)
theme_set(theme_minimal())
library(kableExtra)
options(knitr.table.format = "html") 
library(plotly)
```

My two best friends and I challenged each other to list our top 100 favorite songs of all time at the end of 2017. This is by no means an easy task, but the three of us are just the music-obsessed folks to do it. Still, an all-too-obvious thought came to me as I went about logging my 100 tracks: *What an interesting dataset!* So, once I completed my list, I tried my hand at visualizing it. 

# Data Import

What did my spreadsheet look like? I included the song, artist, album, and year of release right away, but I came up with a couple other ideas as I put my spreadsheet together. 

1. **Genre**: Kind of an obvious one, but I figured it could be interesting. The only thing is, I had to be pretty non-specific when it came to assigning each track a genre. How could I make any reasonable insights from my list if I narrowed in on genre to the point of, say, "emo post-hardcore" or "punk neofolk"? Unfortunately, I had to keep things relatively basic. "Rock" and "alternative" would have to do, as much as it pained me to do it.

2. **Time of discovery**: This is one I'm excited about. I thought about what point in my life I fell in love with each track: childhood, high school, college, or post-grad. Childhood is defined as literally everything before high school, because... well, weren't we all children before high school? 

```{r import}
d <- import("./songs.csv", setclass = "tibble") %>% 
  clean_names()
```

# Year of Release

The first thing I wanted to do was graph my favorite songs by the year they were released. 

```{r release}
ggplot(d, aes(x = year)) +
  geom_density(fill = "cornflower blue", color = "white") +
  labs(x = "Year", y = "Count", title = "Density Distribution of Top 100 Songs", subtitle = "Bimodal distribution reveals contemporary taste with some '70s tendencies") +
  xlim(1950, 2020) +
  theme(plot.title = element_text(size = 30, face = "bold", hjust = 0.5), 
        plot.subtitle = element_text(size = 15, hjust = 0.5),
        legend.title = element_text(size = 20, face = "bold"),
        legend.background = element_rect(size = .5),
        legend.text = element_text(size = 15),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 18, face = "bold"))
```

A few interesting observations:

1. My favorite songs largely come from the last couple of decades, which makes sense. Over the past few years, I've tried to keep up with new music releases, so I'm just listening to more current music than classic. 

2. I almost completely skipped over the '80s. There's just a void for almost that entire decade, with one exception (**"Please, Please, Please Let Me Get What I Want"** by The Smiths). I think that's hilarious. But also, maybe I should listen to more '80s music.

3. I'm going to guess that the entire clump of songs from the '60s and '70s is the influence of my classic rock phase in high school.

# Breaking It Down by Decade

Now that we got a macro view of the entire list by year of release, I thought it would be cool to break things down by decade. I can do that using the `cut` command. Since the earliest decade represented in my list is the '60s, we'll start there and make 10-year slices up until the present decade.

```{r decade_cut}
d <- d %>% 
  mutate(decade = cut(year, breaks = c(1959, 1969, 1979, 1989, 1999, 2009, 2019), 
                      labels = c("'60s", "'70s", "'80s", "'90s", "'00s", "'10s"))) %>% 
  group_by(decade) %>% 
  mutate(decade_count = n())
```

Here is what that first plot looks like when I color the columns differently for each decade represented.

```{r release_decade}
ggplot(d, aes(x = decade, y = decade_count)) +
  geom_bar(stat = "summary", fill = "cornflower blue") +
  geom_text(aes(label = decade_count), vjust = -.5, size = 6, fontface = "bold") +
  labs(x = "Decade", y = "Count") +
  ggtitle("Favorite Songs by Decade of Release", subtitle = "OK, but did the '80s even happen?") +
  theme(plot.title = element_text(size = 30, face = "bold", hjust = 0.5), 
        plot.subtitle = element_text(size = 15, hjust = 0.5),
        legend.title = element_text(size = 20, face = "bold"),
        legend.background = element_rect(size = .5),
        legend.text = element_text(size = 15),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 18, face = "bold"))
```

My tastes are definitely skewed to the contemporary, but the '70s has some nice representation with 15 songs! Also, I've heard some people classify Nirvana as a classic rock band, so I guess even '90s music can be considered "oldies." 

# Genre

Next, I wanted to see my list broken down by genre. Like I mentioned before, the genres aren't nearly as specific as they deserve to be, so take these results with a grain of salt. A genre like "rock," for example, can be broken up so much further. I mean, c'mon, are we talking post-hardcore or proto-metalcore here? `

```{r genre_plot}
d <- d %>% 
  group_by(genre) %>% 
  mutate(genre_freq = n())

ggplot(d, aes(x = reorder(genre, genre_freq), y = genre_freq)) +
  geom_bar(stat = "summary", fill = "cornflower blue", color = "white") +
  geom_text(aes(label = genre_freq), vjust = -.5, size = 6, fontface = "bold") +
  labs(x = "Genre", y = "Count", title = "Favorite Songs by Genre", subtitle = "A frequency distribution that screams 'white dude'") +
  theme(plot.title = element_text(size = 30, face = "bold", hjust = 0.5), 
        plot.subtitle = element_text(size = 15, hjust = 0.5),
        legend.title = element_text(size = 20, face = "bold"),
        legend.background = element_rect(size = .5),
        legend.text = element_text(size = 15),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 18, face = "bold"))
```

Talk about a "diverse" set of music... It looks like 80% of my list is composed of music that can be classified, at least loosely, as rock, alternative, and punk. 

# Period of Discovery

Finally, I wanted to look at the distribution of songs from the period of time in my life I first discovered them. I looked specifically at childhood, high school, college, and post-grad. I'm really excited — and terrified — to see what this one looks like!

```{r discovery}
d$discovery <- factor(d$discovery, levels = c("Childhood", "High School", "College", "Post-Grad"))

d <- d %>% 
  group_by(discovery) %>% 
  mutate(disc_freq = n())

ggplot(d, aes(x = discovery, y = disc_freq)) +
  geom_bar(stat = "summary", fill = "cornflower blue", color = "white") +
  geom_text(aes(label = disc_freq), vjust = -.5, size = 6, fontface = "bold") +
  labs(x = "Period of Discovery", y = "Count", title = "Favorite Songs by Period of Discovery", subtitle = "As the adage goes, college is an important time in one's life") +
  theme(plot.title = element_text(size = 30, face = "bold", hjust = 0.5), 
        plot.subtitle = element_text(size = 15, hjust = 0.5),
        legend.title = element_text(size = 20, face = "bold"),
        legend.background = element_rect(size = .5),
        legend.text = element_text(size = 15),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 18, face = "bold"))
```

It looks like a vast majority (nearly 50%) of my favorite songs came from college. What a formative four years those were... In thinking about it more, I came to realize how obvious this is. Those two best friends I mentioned at the very beginning of this document, the ones who started this year-end list-making tradition with me, I met them my freshman year of college. They know a lot more about music than I ever will, and they introduced me to so many of the songs on this list. My taste in music — and so much more — would not be the same without them.

In sum, you're damn right I discovered most of my picks in college, and with the help of some good friends.

That's revealing enough, but let's see what the genre breakdown looks like for each of the four life phases. Was I really as fixated on classic rock in high school as I suspect I was? Have I broadened my horizons at all since?

```{r discovery_plot2}
d <- d %>% 
  group_by(discovery, genre) %>% 
  mutate(specific = n())

ggplot(d, aes(x = discovery, y = genre, fill = specific)) +
  geom_tile() +
  labs(x = "Period of Discovery", y = "Genre", fill = "Count", title = "Favorite Songs by Genre and Time of Discovery", subtitle = "My tastes have gotten more diverse over time, but only marginally") +
  theme(plot.title = element_text(size = 30, face = "bold", hjust = 0.5), 
        plot.subtitle = element_text(size = 15, hjust = 0.5),
        legend.title = element_text(size = 20, face = "bold"),
        legend.text = element_text(size = 15),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 18, face = "bold"))
```

As this reveals, the answer to both of my hypothetical questions from above is a resounding **YES**. Seriously, 18 entries on my list are rock songs I discovered in high school. I have thankfully diversified my listening in the years since, albeit not by too much. There is a spike in alternative entries from the college years.

# The Ranked List

OK, you've been very patient. After all that, I finally give to you the completed ranked list of my 100 favorite songs, as of 2017. There's even a Spotify playlist embedded in this very document, with all the songs in their correct order for your listening pleasure!

I initially made this document at the end of 2017, after the first term of my PhD program. It's amazing how much I've learned only a year later. I'd never seen myself as a coder or data scientist before grad school, but my program has taught me that anticipated difficulty is no reason to shy away from something new, especially if it's something you care enough about. 

<iframe src="https://open.spotify.com/embed/user/1231312236/playlist/1X0ICzJp8Y6Jt1EWDUP9ot" width="930" height="380" frameborder="10" allowtransparency="true"></iframe>